
 Assume the head and the tail both start at the same position, overlapping.

If the head is ever two steps directly up, down, left, or right from the tail, the tail must also move one step in that direction so it remains close enough

Otherwise, if the head and tail aren't touching and aren't in the same row or column, the tail always moves one step diagonally to keep up.




## Import and process data
 

```{r}
library(tidyverse)
prog <-  read_delim('data/day10_input.txt',delim = ' ', col_names = c('op','arg'))
prog$arg[is.na(prog$arg)] <- 0

prog <- prog |> mutate(
                  steps = case_when(op == 'addx' ~ 2, op == 'noop' ~ 1),
                  step_num_start= tail(accumulate(steps, `+`, .init = 1),-1)-steps,
                  x_reg =  accumulate(arg, `+`), x_reg = x_reg + 1,
                  x_reg_prev = c(0,head(x_reg,-1))
               )
```
 
 
## Part 1
 
Find the signal strength during the 20th, 60th, 100th, 140th, 180th, and 220th cycles. What is the sum of these six signal strengths?
 
 
```{r}
key_steps = seq(20,240,40)

test_key_step <- \(step){
  for(key in key_steps)
    if(step - key == 0 || step - key == -1)
      return(TRUE)
  return(FALSE)
}

prog <- prog |> mutate(key_step = map_lgl(step_num_start, test_key_step)) |> filter(key_step)
sum(key_steps *prog$x_reg_prev)
```

13940 is too high
13200 is too low