 
## Import and process data
 
First thought is to use rows like this:
Money_num, current_items, operation, operand, test_divisor, true_target, false_target

Then we just process these one at a time, updating the tibble with new current items after each item is processed. Or might be better to keep a separate list of current items to avoid all the copying?


```{r}
library(tidyverse)
raw_data <-  read_file('data/day11_input.txt')
```
 
I assume there is a much better way!

```{r}
setup <- \(){
  monkey_num <<- as.numeric(str_match_all(raw_data,"Monkey (\\d+)")[[1]][,2])
  operation  <<- str_match_all(raw_data,"Operation: new = old (.)")[[1]][,2]
  operand    <<- str_match_all(raw_data,"Operation: new = old . (.+)")[[1]][,2]
  test_divisor <<- as.numeric(str_match_all(raw_data,"Test: divisible by (\\d+)")[[1]][,2])
  true_target <<- as.numeric(str_match_all(raw_data,"If true: throw to monkey (\\d+)")[[1]][,2])
  false_target <<- as.numeric(str_match_all(raw_data,"If false: throw to monkey (\\d+)")[[1]][,2])
  items <<- str_match_all(raw_data,"Starting items: ((\\d+, )*(\\d+))") [[1]][,2]
  items <<- map(items, \(x){as.numeric(str_split(x,", ")[[1]])})
}

```


## Part 1



```{r}
inspect_item <- \(monkey, item){
  if(operand[[monkey]] == "old")
      value <- item
  else
      value <- as.numeric(operand[[monkey]])
  
  if(operation[[monkey]] == "*")
     return(item * value)
  return(item + value)
}

```

* Assumes (as in my data) that monkeys never throw to themselves

```{r}
setup()
num_rounds = 20
monkey_inspections <- rep(0,length(monkey_num))
for(round in 1:num_rounds){
  for(monkey_index in seq_along(monkey_num)){
    curr_items <- items[[monkey_index]]
    monkey_inspections[[monkey_index]] <- monkey_inspections[[monkey_index]] + length(curr_items)
    if(length(curr_items) > 0) {
      for(i in seq_along(curr_items)){
         final_worry <-  inspect_item(monkey_index, curr_items[[i]]) %/% 3
         if(final_worry %% test_divisor[[monkey_index]] == 0){
           target <- true_target[[monkey_index]] + 1
         } else {
           target <- false_target[[monkey_index]] + 1
         }
         items[[ target ]] <- c(items[[target]], final_worry)
      }
      items[[monkey_index]] <- double()  #empty 
    }
  }
}
prod(tail(sort(monkey_inspections),2))    
```

 
113232


## Part 2 

NOTE : Need to set up to reload data arrays

```{r}
setup()
num_rounds = 10000
reducer = prod(test_divisor)
monkey_inspections <- rep(0,length(monkey_num))
for(round in 1:num_rounds){
  for(monkey_index in seq_along(monkey_num)){
    curr_items <- items[[monkey_index]]
    monkey_inspections[[monkey_index]] <- monkey_inspections[[monkey_index]] + length(curr_items)
    if(length(curr_items) > 0) {
      for(i in seq_along(curr_items)){
         final_worry <-  inspect_item(monkey_index, curr_items[[i]]) %% reducer
         if(final_worry %% test_divisor[[monkey_index]] == 0){
           target <- true_target[[monkey_index]] + 1
         } else {
           target <- false_target[[monkey_index]] + 1
         }
         items[[ target ]] <- c(items[[target]], final_worry)
      }
      items[[monkey_index]] <- double()  #empty 
    }
  }
}
prod(tail(sort(monkey_inspections),2))
```

 


 